Tool-driven Revolutions in Archaeological Science

### 2018-09-11 analysis/paper/jcaa-paper.Rmd
corrected citation
paradigm to just change
Yeah, we should add some commentary in the end why exactly we think so.

In this paper we investigate recent developments which facilitate computational archaeology to return to principles of the scientific method and whether this might be the beginning of a paradigm-shift. Revolutions or paradigm-shifts in archaeology have been discussed to happen following theoretical discourses [@Clark_1993; @Härke_2002], but there has been an argument in philosophy of science whether revolutions in science might not just be ideas-driven but also tool-driven [@Galison_1997; @Dyson_2000]. In this light we explore how Ecology and Archaeology have started to embrace methods which improve the reproducibility of their research. We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages, exemplified by R, and the sharing of code. A review of bibliometric evidence showcases a strong development towards the use of R for Ecology and hints at the beginning of a similar development in archaeology. We thus evaluate how approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and propose these might be the sign of a tool-driven paradigm-change in archaeology. Acknowledging this is not a simple process, we offer a R-based tool to facilitates the task of creating a compendium which enables other researchers to reproduce the published results.

@@

Nonetheless we feel that with the introduction of computers there has been a gradual and unintended drift away from reproducibility, and thus from scientific practice. In the digital age, many archaeological tasks have shifted from the descriptive and qualitative aspects to quantitative and computational analyses. This shift also mandates a change in documentation of analyses to enable other researchers to reproduce published results. Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which obscures and complicates the reproducibility of analyses. By using and sharing code we can undo this drift, and return to a robust scrietific practice for archaeology. As we show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with natural and physical sciences. These sciences also show a trend to increased reproducibility as Journals change their data and code access policies, demanding code and data to be published alongside articles [@Stodden_2013]. Also in archaeology, more and more funding agencies push for open science methods and transparency [@Stupler_2017].

>>>>

In this paper we investigate recent developments which facilitate computational archaeology to return to principles of the scientific method and whether this might be the beginning of a paradigm-shift. Revolutions or paradigm-shifts in archaeology have been discussed to happen following theoretical discourses [@Clark_1993; @Härke_2002], but there has been an argument in philosophy of science whether revolutions in science might not just be ideas-driven but also tool-driven [@Galison_1997; @Dyson_2000]. In this light we explore how Ecology and Archaeology have started to embrace methods which improve the reproducibility of their research. We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages, exemplified by R, and the sharing of code. A review of bibliometric evidence showcases a strong development towards the use of R for Ecology and hints at the beginning of a similar development in archaeology. We thus evaluate how approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and propose these might be the sign of a tool-driven change in archaeology. Acknowledging this is not a simple process, we offer a R-based tool to facilitates the task of creating a compendium which enables other researchers to reproduce the published results.

@@

Nonetheless we feel that with the introduction of computers there has been a gradual and unintended drift away from reproducibility, and thus from scientific practice. In the digital age, many archaeological tasks have shifted from the descriptive and qualitative aspects to quantitative and computational analyses. This shift also mandates a change in documentation of analyses to enable other researchers to reproduce published results. Modern archaeology has often been lacking in this aspect, as the research has been undertaken using point-and-click software, which obscures and complicates the reproducibility of analyses. By using and sharing code we can undo this drift, and return to a robust scrietific practice for archaeology. As we show, the advance of scripting languages takes place mostly in the part of archaeology that is in close contact with natural and physical sciences. These sciences also show a trend to increased reproducibility as Journals change their data and code access policies, demanding code and data to be published alongside articles [@Stodden_2013]. Also in archaeology, more and more funding agencies push for open science methods and transparency [@Strupler_2017].
